---
title: "Installing TensorFlow"
output: 
  html_document:
    toc_depth: 3
    toc_float:
        collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Overview

Prior to using the **tensorflow** R package you need to install a version of TensorFlow on your system. Below we describe how to install TensorFlow as well the various options available for customizing your installation. 

Note that this article principally covers the use of the R `install_tensorflow()` function, which provides an easy to use wrapper for the various steps required to install TensorFlow.

You can also choose to install TensorFlow manually (as described at <https://www.tensorflow.org/install/>). In that case the [Custom Installation] section covers how to arrange for the tensorflow R package to use the version you installed.

## Quick Start

First, install the tensorflow R package from GitHub as follows:

```{r}
devtools::install_github("rstudio/tensorflow")
```

Then, use the `install_tensorflow()` function to install TensorFlow:

```{r}
library(tensorflow)
install_tensorflow()
```

You can confirm that the installation succeeded with:

```{r}
sess = tf$Session()
hello <- tf$constant('Hello, TensorFlow!')
sess$run(hello)
```

This will provide you with a default installation of TensorFlow suitable for use with the tensorflow R package. Read on if you want to learn about additional installation options.

## Installation Methods

TensorFlow is distributed as a Python package and so needs to be installed within a Python environment on your system. By default, the `install_tensorflow()` function attempts to install TensorFlow within it's own isolated Python environment ("r-tensorflow").

These are the available methods and their behavior:

| Method  | Description  |
|--------------|----------------------------------------------------------|
| auto  | Automatically choose an appropriate default for the current platform. |
| virtualenv  | Install into a Python virtual environment at `~/.virtualenvs/r-tensorflow` |
| conda  | Install into an Anaconda Python environment named `r-tensorflow` |
| system  | Install into the system Python environment  |

The "virtualenv" and "conda" methods are available on Linux and OS X and the "conda" and "system" methods are available on Windows. To override the default behavior use the `method` parameter, for example:

```{r}
install_tensorflow(method = "conda")
```

## CPU vs. GPU

TensorFlow can be configured to run on either CPUs or GPUs. The CPU version is much easier to install and configure so is the best starting place especially when you are first learning how to use TensorFlow. Here's the guidance on CPU vs. GPU versions from the TensorFlow website:

- **TensorFlow with CPU support only**. If your system does not have a NVIDIA® GPU, you must install this version. Note that this version of TensorFlow is typically much easier to install (typically, in 5 or 10 minutes), so even if you have an NVIDIA GPU, we recommend installing this version first.

- **TensorFlow with GPU support**. TensorFlow programs typically run significantly faster on a GPU than on a CPU. Therefore, if your system has a NVIDIA® GPU meeting the prerequisites shown below and you need to run performance-critical applications, you should ultimately install this version.

By default `install_tensorflow()` installs the CPU version of TensorFlow. If you want to use the GPU version you should first ensure that your system meets all of these prerequisites:

- Running an NVIDIA® GPU

- Has the various CUDA libraries correctly installed and configured. This varies by platform, see:

    - [Linux](https://www.tensorflow.org/install/install_linux#nvidia_requirements_to_run_tensorflow_with_gpu_support)
    - [OS X](https://www.tensorflow.org/install/install_mac#requirements_to_run_tensorflow_with_gpu_support)
    - [Windows](https://www.tensorflow.org/install/install_windows#requirements_to_run_tensorflow_with_gpu_support)

Once you meet these prerequisites you can install TensorFlow with GPU support via:

```{r}
install_tensorflow(gpu = TRUE)
```

## Alternate Versions

By default, `install_tensorflow()` install the latest release version of TensorFlow. You can override this behavior by specifying the `version` parameter. For example:

```{r}
install_tensorflow(version = "1.0.0")
```

Note that this should be a full `major.minor.patch` version specification (rather than just major and minor versions).

You can also install nightly or other special builds of TensorFlow by specifying an explicit `package_url` parameter. For example:

```{r}
install_tensorflow(package_url = "https://ci.tensorflow.org/view/Nightly/job/...")
```

Note that if you specify an explicit `package_url` this will override the `version` and `gpu` parameters and install whatever version is referenced by the URL.

You can find links to the nightly TensorFlow builds [here](https://github.com/tensorflow/tensorflow#installation).

## Custom Installation

The `install_tensorflow()` function is provided as a convenient way to get started, but is not required. If you have an existing installation of TensorFlow or just prefer your own custom installation that's fine too.

The full instructions for installing TensorFlow on various platforms are here: <https://www.tensorflow.org/install/>. After installing, please refer to the sections below on locating TensorFlow and meeting additional dependencies to ensure that the tensorflow for R package functions correctly with your installation.

### Locating TensorFlow

Once you've installed TensorFlow you need to ensure that the tensorflow for R package can find your installation. The package scans the system for various versions of Python, and also scans available virtual environments and conda environments, so in many cases things will just work without additional effort. 

If the version of TensorFlow you installed is not found automatically, then you can use the following techniques to ensure that TensorFlow is located.

Specify the `TENSORFLOW_PYTHON` environment variable to force probing within a specific Python installation. For example:

```{r}
library(tensorflow)
Sys.setenv(TENSORFLOW_PYTHON="/usr/local/bin/python")
```

Alternatively, call the `use_python` family of configuration functions:

| Function  | Description | 
|----------------|------------------------------------------------------------------|
| `use_python()` | Specify the path a specific Python binary. | 
| `use_virtualenv()` | Specify the directory containing a Python virtualenv. | 
| `use_condaenv()` | Specify the name of a conda environment. | 

For example:

```{r}
library(tensorflow)
use_python("/usr/local/bin/python")
use_virtualenv("~/myenv")
use_condaenv("myenv")
```

Note that you can include multiple calls to the `use_` functions and all provided locations will be tried in the order they were specified.

### Additional Dependencies

There are some components of TensorFlow (e.g. the [Keras](https://github.com/tensorflow/tensorflow/tree/master/tensorflow/contrib/keras) library) which have dependencies on additional Python packages. The `install_tensorflow()` function installs these dependencies automatically, however if you do a custom installation you should be sure to install them manually. 

You can install the additional dependencies with the following command:

```bash
pip install h5py pyyaml requests Pillow scipy
```









